### 2018-03-27 paper/sexSelectionSpacing-ver1.tex
New abstract and minor edits

Strong son preference has frequently been associated with shorter 
spacing between births after the birth of a girl.
The spread of access to prenatal sex determination and 
sex-selective abortions has the potential to reverse this pattern
because each abortion adds substantially to the duration between births.
I introduce a statistical method that simultaneously account for how sex 
selection increases the likelihood of a son \emph{and} the spacing between births.
Using India's National Family and Health Surveys,
I show that use of sex selection leads to \emph{longer} spacing after a daughter 
than after a son; a reversal of the pattern before the availability of sex selection.
For families that do not use sex selection, the duration to next birth is still 
shorter when no boys are present.

Women with 8 or more years of education, both in urban and rural areas, are 
the main users of sex-selective abortions, whereas women with less education 
do not appear to use sex selection.

@@

% [Results]
There are four main results.
First, there has been a general increase in the length of spacing between births
over the four decades covered by the data.
The exception is for first births, where the median duration has either remained
the same or slightly declined, although this hides a substantial compression of
the variation in spell length.
Second, consistent with the use of sex selection, the increase in duration is 
substantially longer for women who are most likely to use sex selection than
for those less likely to use sex selection.
Specifically, for the best educated women, those with no boys now has 
significantly longer spacing and a more male-biased sex ratio than those 
with boys.
Third, women with no education still follow the standard pattern of short 
spacing when they have girls and little evidence of the use of sex selection.
In other word, these women follow a strategy where they achieve a son through
higher fertility rather than use of sex selection.
Finally, sex ratios are more likely to decline within spells at lower parities, 
where the pressure to ensure a son is lower, and more likely to increase or 
remain consistently high for higher-order spells, where the pressure to ensure 
a son is high.

>>>>

Strong son preference is typically associated with shorter birth spacing
in the absence of sons, but access to sex selection has the potential to
reverse this pattern because each abortion extends spacing by six to
twelve months. 
I introduce a statistical method that simultaneously
accounts for how sex selection increases both the likelihood of a son
and the spacing between births. 
Using four rounds of India's National
Family and Health Surveys, I first show that, except for first births,
the spacing between births increased substantially over the last four
decades. 
Birth spacing increased the most for women most likely to use
sex selection; specifically, well-educated women with no boys now
exhibit significantly longer spacing and more male-biased sex ratios
than those with boys. 
Women with no education still follow the standard
pattern of short spacing when they have girls and little evidence of sex
selection, with medium-educated women showing mixed results. 
Finally,
sex ratios are more likely to decline within spells at lower parities,
where there is less pressure to ensure a son, and more likely to
increase or remain consistently high for higher-order spells, where the
pressure to provide a son is high.

@@

% [Results]
There are four main results.
First, there has been a general increase in the length of spacing between births
over the four decades covered by the data.
The exception is for first births, where the median duration has either remained
the same or slightly declined, although this hides a substantial compression of
the variation in spell length.
Second, the increase in duration is substantially larger for the groups of women 
who are most likely to use sex selection.
Specifically, among the best educated women, those with no boys now has 
significantly longer spacing and a more male-biased sex ratio than those 
with boys.
Third, women with no education still follow the standard pattern of short 
spacing when they have girls and little evidence of the use of sex selection.
In other word, these women follow a strategy where they achieve a son through
higher fertility rather than use of sex selection.
Finally, sex ratios are more likely to decline within spells at lower parities, 
where the pressure to ensure a son is lower, and more likely to increase or 
remain consistently high for higher-order spells, where the pressure to ensure 
a son is high.
